The K2 Halo Photometry Campaign

### 2019-04-05 paper/ms.tex
updates

% describe saturation

% describe halo algorithm
We use an extension of the method of halo photometry, first described by \citet{White2017} and with further developments by \citet{Farr2018}. Because SAP is not possible, we consider instead the unsaturated pixels $p_j$ at the wings of the broad and complex PSF, and construct a light curve as a linear combination with weights $w_j$ of these time series, so that flux at cadence $i$ is

@@

\noindent The weights are chosen to minimize an objective function

@@

\noindent which is the $L_k$ norm on a `lagged' finite difference with a lag parameter $l$. For $k = 1$ and $l = 1$, $\Lambda_{1,1}$ is the standard Total Variation as used in previous halo papers, and can be seen as the L1 norm on the derivative of $f$ or as a discrete approximation to its arc length. The lag parameter $l$ allow for flexibility in modelling systematics occurring at different timescales from cadence-to-cadence, and we investigate its effects below. The order parameter $k$ allows for flexibility in how sensitive we are to normally-distributed versus long-tailed noise. TV is a convex objective function \textcolor{red}{and I think the generalizations are too} and $\Lambda_{k,l}$ has analytic derivatives with respect to $w_j$ \citep[calculated with \texttt{autograd}][]{autograd}, and it is therefore extremely fast to optimize and converges well on a global solution.

@@

In order to choose the objective function $\Lambda_{k,l}$, we have selected the quiet star 36~Ophiuchi, the lowest-mass main sequence star in the halo sample and one with very little high frequency variability detected or predicted, which was also observed at short cadence. We choose the 6.5~hour Combined Differential Photometric Precision \citep[CDPP,][]{cdpp} as implemented in \texttt{lightkurve} \citep{lightkurve} as a proxy for the `noise' in a lightcurve, with lower being better. We calculate halo lightcurves and their CDPPs for $k \in \{1,2,3\}$, and $l \in [1,50]$ for long cadence and (sparsely) $l \in [1,2500]$ for short cadence data. The results are displayed in Figure~\ref{cdpps}. We find that for long cadence data, the L1 norm (TV) and a lag of 10 provide the best CDPP, which is unsurprising given that this is a slightly shorter lag than the 12~cadence thruster firing period. In this context we can understand the optimum as TV-min suppressing systematics on the same timescale as they occur. On the other hand, for short cadence data, performance at short lags is very poor but the method performs similarly for all values of $k$ and improves very slowly with $l$ for longer lags. In calculating lightcurves here, seeing the very weak dependence of short cadence CDPP on $l$, we calculate our final light curves with TV and a lag of 300 cadences for consistency with the long cadence method.

@@

% link to tables 
% normal apertures in 6

The full sample of stars for which halo apertures were obtained is listed in Tables~\ref{table_0}--\ref{table_5}, broken down by Campaign. While some very bright stars were observed with conventional apertures as part of these programs, simple aperture photometry is satisfactory on these targets and we exclude them from the present discussion and data release, which is oriented strictly towards targets only observable with halo photometry. We make an exception for Spica, which was observed in Campaign~6 without a halo aperture but in Campaign~17 with a halo aperture. In Campaign~6 it was assigned a normal aperture by mistake and simple aperture photometry performed extremely poorly, so we have processed it with the halo pipeline. The stars in Campaign~18 were also on-silicon in Campaign~5, but were not assigned apertures suitable for halo photometry in C5. A possible further extension of the present work would be to recover C5 light curves for these objects using either or both of smear or modified halo photometry.

@@

Some of the objects here have been previously published, but we here provide the first public data releases for the Pleiades' Seven Sisters \citep{White2017}, Aldebaran \citep{Farr2018}, $\iota$~Lib \citep{Buysschaert2018}, and $\epsilon$~Tau \citep{Arentoft2019}, as well as $\rho$~Leo with halo photometry but without using the TV-min method \citep{Aerts2018}.

>>>>

% describe saturation
The \kepler detector saturates at a magnitude of $K_p \sim 11.3$ in both long- (30~min) and short (1~min)-cadence data, as these both represent sums of 6~s exposures \citep{Gilliland2010}. Beyond this point, additional photons cause excess electrons to `bleed' into adjacent pixels in both directions along the column containing the bright star. This process occurs conservatively and simple aperture photometry (SAP) adding all the flux contained in a window around the bleed column has recovered light curves with precisions close to the photon noise limit of stars as bright as 16~Cyg~AB, $\theta$~Cyg and RR~Lyr \citep[e.g.][]{Kolenberg2011,2013MNRAS.433.1262W,Guzik2016}. In the nominal \kepler mission this was not attempted for all such bright stars, and in \ktwo, the several-pixel spacecraft motion significantly increased the size of the required apertures for SAP photometry of very saturated stars, while also making their systematics more difficult to deal with. While the second-version pixel-level-decorrelation (PLD) pipeline EVEREST~2.0 was able to correct systematics in saturated SAP photometry \citep{everest2}, this is rendered impossible for the very brightest stars whose bleed columns may run to the edge of the detector. Furthermore, bandwidth constraints meant that pixel data were not downloaded for many bright targets in \ktwo. It was therefore necessary to develop a method to recover good light curves from a smaller number of unsaturated pixels in a compact aperture around bright stars: halo photometry.

% describe halo algorithm
This method was first described by \citet{White2017} and applied to the Pleiades' Seven Sisters, and was applied to Aldebaran with further developments by \citet{Farr2018}. Because SAP is not possible, we consider instead the unsaturated pixels $p_j$ at the wings of the broad and structured PSF. We construct a light curve as a linear combination of these time series with weights $w_j$, so that flux at cadence $i$ is

@@

\noindent In our updated pipeline presented here, the weights are chosen to minimize an objective function

@@

\noindent which is the $L_k$ norm on a `lagged' finite difference with a lag parameter $l$. For $k = 1$ and $l = 1$, $\Lambda_{1,1}$ is the standard Total Variation objective (TV) used in previous halo papers, and can be seen as the L1 norm on the derivative of $f$ or as a discrete approximation to its arc length. The lag parameter $l$ allows for flexibility in modelling systematics occurring at different timescales from cadence-to-cadence, and we investigate its effects below. The order parameter $k$ allows for flexibility in how sensitive we are to normally-distributed versus long-tailed noise. TV is a convex objective function \textcolor{red}{and I think the generalizations are too} and $\Lambda_{k,l}$ has analytic derivatives with respect to $w_j$ \citep[calculated with \texttt{autograd}][]{autograd}, and it is therefore extremely fast to optimize and converges well on a global solution.

@@

In order to choose the objective function $\Lambda_{k,l}$, we have selected the quiet star 36~Ophiuchi, the lowest-mass main sequence star in the halo sample and one with very little high frequency variability detected or predicted, which was also observed at short cadence. We choose the 6.5~hour Combined Differential Photometric Precision \citep[CDPP,][]{cdpp} as implemented in \texttt{lightkurve} \citep{lightkurve} as a proxy for the `noise' in a lightcurve, with lower being better. We calculate halo lightcurves and their CDPPs for $k \in \{1,2,3\}$, and $l \in [1,50]$ for long cadence and (sparsely) $l \in [1,2500]$ for short cadence data. The results are displayed in Figure~\ref{cdpps}. We find that for long cadence data, the L1 norm (TV) and a lag of 10 provide the best CDPP, which is unsurprising given that this is a slightly shorter lag than the 12~cadence thruster firing period. In this context we can understand the optimum as TV-min suppressing systematics on the same timescale as they occur. On the other hand, for short cadence data, performance at short lags is very poor but the method performs similarly for all values of $k$ and improves very slowly with $l$ for longer lags.

@@

% link to tables

The full sample of stars for which halo apertures were obtained is listed in Tables~\ref{table_0}--\ref{table_5}, broken down by Campaign. While some very bright stars were observed with conventional apertures as part of these programs, simple aperture photometry is satisfactory on these targets and we exclude them from the present discussion and data release, which is oriented strictly towards targets only observable with halo photometry. 
% normal apertures in 6
We make an exception for Spica, which was observed in Campaign~6 without a halo aperture but in Campaign~17 with a halo aperture. In Campaign~6 it was assigned a normal aperture by mistake and simple aperture photometry performed extremely poorly, so we have processed it with the halo pipeline. The stars in Campaign~18 were also on-silicon in Campaign~5, but were not assigned apertures suitable for halo photometry in C5. A possible further extension of the present work would be to recover C5 light curves for these objects using either or both of smear or modified halo photometry.

@@

Analyses for several of the objects here have been previously published, but we here provide the first public data releases for the Pleiades' Seven Sisters \citep{White2017}, Aldebaran \citep{Farr2018}, $\iota$~Lib \citep{Buysschaert2018}, and $\epsilon$~Tau \citep{Arentoft2019}, as well as $\rho$~Leo which was studied with halo pixels but without our objective functions \citep{Aerts2018}.
